---
title: "Boosting with regularization"
author: "Cameron Roach"
output:
  pdf_document:
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	cache.path = "./../cache/knitr/"
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
require(plotly)
require(DT)
require(caret)
require(myhelpr)
require(doMC)

registerDoMC(cores = 3)

source("../R/loadData.R")
source("../R/clean-smd-data.R")

load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)

smd <- load_smd_data(load_zones, root_dir = "./..")
smd <- clean_smd_data(smd, root_dir = "./..")
```

# Introduction

Here I wish to test the impact of a L1 and L2 regularization on model accuracy when using boosting.

Electricity demand in the ME load zone will be used. Only working days will be considered (weekends and holidays are filtered). Dry bulb temperature and lags will be used as a predictors. Models will be assessed against each other using RMSE.

## Lagged variable creation

```{r gen_dummy_vars}
iZ <- "ME"

smd <- smd %>%
  filter(Zone == iZ,
         Weekend == FALSE,
         Holiday_flag == FALSE) %>% 
  mutate(DryBulb_lag1 = lag(DryBulb, 1),
         DryBulb_lag2 = lag(DryBulb, 2),
         DryBulb_lag3 = lag(DryBulb, 3),
         DryBulb_lag4 = lag(DryBulb, 4),
         DryBulb_lag5 = lag(DryBulb, 5),
         DryBulb_lag6 = lag(DryBulb, 6),
         DryBulb_lag7 = lag(DryBulb, 7),
         DryBulb_lag8 = lag(DryBulb, 8),
         DryBulb_lag9 = lag(DryBulb, 9),
         DryBulb_lag10 = lag(DryBulb, 10),
         DryBulb_lag11 = lag(DryBulb, 11),
         DryBulb_lag12 = lag(DryBulb, 12),
         DryBulb_lag13 = lag(DryBulb, 13),
         DryBulb_lag14 = lag(DryBulb, 14),
         DryBulb_lag15 = lag(DryBulb, 15),
         DryBulb_lag16 = lag(DryBulb, 16),
         DryBulb_lag17 = lag(DryBulb, 17),
         DryBulb_lag18 = lag(DryBulb, 18),
         DryBulb_lag19 = lag(DryBulb, 19),
         DryBulb_lag20 = lag(DryBulb, 20),
         DryBulb_lag21 = lag(DryBulb, 21),
         DryBulb_lag22 = lag(DryBulb, 22),
         DryBulb_lag23 = lag(DryBulb, 23),
         DryBulb_lag24 = lag(DryBulb, 24),
         DryBulb_lag25 = lag(DryBulb, 25),
         DryBulb_lag26 = lag(DryBulb, 26),
         DryBulb_lag27 = lag(DryBulb, 27),
         DryBulb_lag28 = lag(DryBulb, 28),
         DryBulb_lag29 = lag(DryBulb, 29),
         DryBulb_lag30 = lag(DryBulb, 30),
         DryBulb_lag31 = lag(DryBulb, 31),
         DryBulb_lag32 = lag(DryBulb, 32),
         DryBulb_lag33 = lag(DryBulb, 33),
         DryBulb_lag34 = lag(DryBulb, 34),
         DryBulb_lag35 = lag(DryBulb, 35),
         DryBulb_lag36 = lag(DryBulb, 36),
         DryBulb_lag37 = lag(DryBulb, 37),
         DryBulb_lag38 = lag(DryBulb, 38),
         DryBulb_lag39 = lag(DryBulb, 39),
         DryBulb_lag40 = lag(DryBulb, 40),
         DryBulb_lag41 = lag(DryBulb, 41),
         DryBulb_lag42 = lag(DryBulb, 42),
         DryBulb_lag43 = lag(DryBulb, 43),
         DryBulb_lag44 = lag(DryBulb, 44),
         DryBulb_lag45 = lag(DryBulb, 45),
         DryBulb_lag46 = lag(DryBulb, 46),
         DryBulb_lag47 = lag(DryBulb, 47),
         DryBulb_lag48 = lag(DryBulb, 48),
         DryBulb_lag49 = lag(DryBulb, 49),
         DryBulb_lag50 = lag(DryBulb, 50),
         DryBulb_lag51 = lag(DryBulb, 51),
         DryBulb_lag52 = lag(DryBulb, 52),
         DryBulb_lag53 = lag(DryBulb, 53),
         DryBulb_lag54 = lag(DryBulb, 54),
         DryBulb_lag55 = lag(DryBulb, 55),
         DryBulb_lag56 = lag(DryBulb, 56),
         DryBulb_lag57 = lag(DryBulb, 57),
         DryBulb_lag58 = lag(DryBulb, 58),
         DryBulb_lag59 = lag(DryBulb, 59),
         DryBulb_lag60 = lag(DryBulb, 60),
         DryBulb_lag61 = lag(DryBulb, 61),
         DryBulb_lag62 = lag(DryBulb, 62),
         DryBulb_lag63 = lag(DryBulb, 63),
         DryBulb_lag64 = lag(DryBulb, 64),
         DryBulb_lag65 = lag(DryBulb, 65),
         DryBulb_lag66 = lag(DryBulb, 66),
         DryBulb_lag67 = lag(DryBulb, 67),
         DryBulb_lag68 = lag(DryBulb, 68),
         DryBulb_lag69 = lag(DryBulb, 69),
         DryBulb_lag70 = lag(DryBulb, 70),
         DryBulb_lag71 = lag(DryBulb, 71),
         DryBulb_lag72 = lag(DryBulb, 72),
         DewPnt_lag1 = lag(DewPnt, 1),
         DewPnt_lag2 = lag(DewPnt, 2),
         DewPnt_lag3 = lag(DewPnt, 3),
         DewPnt_lag4 = lag(DewPnt, 4),
         DewPnt_lag5 = lag(DewPnt, 5),
         DewPnt_lag6 = lag(DewPnt, 6),
         DewPnt_lag7 = lag(DewPnt, 7),
         DewPnt_lag8 = lag(DewPnt, 8),
         DewPnt_lag9 = lag(DewPnt, 9),
         DewPnt_lag10 = lag(DewPnt, 10),
         DewPnt_lag11 = lag(DewPnt, 11),
         DewPnt_lag12 = lag(DewPnt, 12),
         DewPnt_lag13 = lag(DewPnt, 13),
         DewPnt_lag14 = lag(DewPnt, 14),
         DewPnt_lag15 = lag(DewPnt, 15),
         DewPnt_lag16 = lag(DewPnt, 16),
         DewPnt_lag17 = lag(DewPnt, 17),
         DewPnt_lag18 = lag(DewPnt, 18),
         DewPnt_lag19 = lag(DewPnt, 19),
         DewPnt_lag20 = lag(DewPnt, 20),
         DewPnt_lag21 = lag(DewPnt, 21),
         DewPnt_lag22 = lag(DewPnt, 22),
         DewPnt_lag23 = lag(DewPnt, 23),
         DewPnt_lag24 = lag(DewPnt, 24),
         DewPnt_lag25 = lag(DewPnt, 25),
         DewPnt_lag26 = lag(DewPnt, 26),
         DewPnt_lag27 = lag(DewPnt, 27),
         DewPnt_lag28 = lag(DewPnt, 28),
         DewPnt_lag29 = lag(DewPnt, 29),
         DewPnt_lag30 = lag(DewPnt, 30),
         DewPnt_lag31 = lag(DewPnt, 31),
         DewPnt_lag32 = lag(DewPnt, 32),
         DewPnt_lag33 = lag(DewPnt, 33),
         DewPnt_lag34 = lag(DewPnt, 34),
         DewPnt_lag35 = lag(DewPnt, 35),
         DewPnt_lag36 = lag(DewPnt, 36),
         DewPnt_lag37 = lag(DewPnt, 37),
         DewPnt_lag38 = lag(DewPnt, 38),
         DewPnt_lag39 = lag(DewPnt, 39),
         DewPnt_lag40 = lag(DewPnt, 40),
         DewPnt_lag41 = lag(DewPnt, 41),
         DewPnt_lag42 = lag(DewPnt, 42),
         DewPnt_lag43 = lag(DewPnt, 43),
         DewPnt_lag44 = lag(DewPnt, 44),
         DewPnt_lag45 = lag(DewPnt, 45),
         DewPnt_lag46 = lag(DewPnt, 46),
         DewPnt_lag47 = lag(DewPnt, 47),
         DewPnt_lag48 = lag(DewPnt, 48),
         DewPnt_lag49 = lag(DewPnt, 49),
         DewPnt_lag50 = lag(DewPnt, 50),
         DewPnt_lag51 = lag(DewPnt, 51),
         DewPnt_lag52 = lag(DewPnt, 52),
         DewPnt_lag53 = lag(DewPnt, 53),
         DewPnt_lag54 = lag(DewPnt, 54),
         DewPnt_lag55 = lag(DewPnt, 55),
         DewPnt_lag56 = lag(DewPnt, 56),
         DewPnt_lag57 = lag(DewPnt, 57),
         DewPnt_lag58 = lag(DewPnt, 58),
         DewPnt_lag59 = lag(DewPnt, 59),
         DewPnt_lag60 = lag(DewPnt, 60),
         DewPnt_lag61 = lag(DewPnt, 61),
         DewPnt_lag62 = lag(DewPnt, 62),
         DewPnt_lag63 = lag(DewPnt, 63),
         DewPnt_lag64 = lag(DewPnt, 64),
         DewPnt_lag65 = lag(DewPnt, 65),
         DewPnt_lag66 = lag(DewPnt, 66),
         DewPnt_lag67 = lag(DewPnt, 67),
         DewPnt_lag68 = lag(DewPnt, 68),
         DewPnt_lag69 = lag(DewPnt, 69),
         DewPnt_lag70 = lag(DewPnt, 70),
         DewPnt_lag71 = lag(DewPnt, 71),
         DewPnt_lag72 = lag(DewPnt, 72))

train_data <- filter(smd, Year >= 2012, Year < 2015)

test_data <- filter(smd, Year == 2015)
```


# Boosting with L1-regularization

We will use all predictors and test a numer of weights for L1 regularization. All predictors (temperatures and lagged temperatures) have been scaled except for period of day, which is treated as a categorical variable.

```{r fit_boost_l1_reg}
xgb_ctrl <- trainControl(method = "repeatedcv",
                         number = 5,
                         allowParallel = TRUE)

xgb_grid_linear <- expand.grid(nrounds = 300,
                               alpha = 10^((0:8)/2),
                               lambda = 0,
                               eta = 0.1)

system.time({
  xgb_l1_fit <- train_data %>% 
    select(Demand, Period, starts_with("DryBulb"), starts_with("DewPnt")) %>% 
    mutate_at(vars(starts_with("DryBulb"), starts_with("DewPnt")),
              scale, center = FALSE) %>% 
    train(Demand ~ . ,
          data = .,
          method="xgbLinear",
          trControl = xgb_ctrl,
          tuneGrid = xgb_grid_linear,
          nthread = 1) # avoids OpenMP and doMC clash
})
```

## L1-regularization results

We see lowest RMSE's when `alpha` is set to 1000.

```{r model_performance_l1}
xgb_l1_fit

ggplot(xgb_l1_fit) +
  scale_x_log10() +
  ggtitle("L1 model performance")
```



# Boosting with L2-regularization

We will use all predictors and test a numer of weights for L2 regularization. All predictors (temperatures and lagged temperatures) have been scaled except for period of day, which is treated as a categorical variable.

```{r fit_boost_l2_reg}
xgb_ctrl <- trainControl(method = "repeatedcv",
                         number = 5,
                         allowParallel = TRUE)

xgb_grid_linear <- expand.grid(nrounds = 300,
                               alpha = 0,
                               lambda = 10^((0:8)/2),
                               eta = 0.1)

system.time({
  xgb_l2_fit <- train_data %>% 
    select(Demand, Period, starts_with("DryBulb"), starts_with("DewPnt")) %>% 
    mutate_at(vars(starts_with("DryBulb"), starts_with("DewPnt")),
              scale, center = FALSE) %>% 
    train(Demand ~ . ,
          data = .,
          method="xgbLinear",
          trControl = xgb_ctrl,
          tuneGrid = xgb_grid_linear,
          nthread = 1)
})
```

## L2-regularization results

We see lowest RMSE's when `lambda` is set to 100. It looks as though the L2-regularization produces significantly better results than L1. This is most likely due to the strong correlation in the predictor variables. L1 regularization will give quite different values to correlated variables, whereas L2 regularization gives similar values (see ISLR p. 261).

```{r model_performance_l2}
xgb_l2_fit

ggplot(xgb_l2_fit) +
  scale_x_log10() +
  ggtitle("L2 model performance")
```

# Test data results

Let's calculate the RMSE score for the fitted model on the test data set. This will give a better idea of how well it really performs.

```{r test_performance}
scaled_test_data <-  test_data %>% 
    select(Demand, Period, starts_with("DryBulb"), starts_with("DewPnt")) %>% 
    mutate_at(vars(starts_with("DryBulb"), starts_with("DewPnt")),
              scale, center = FALSE)

test_data$Demand_pred_l1 <- predict(xgb_l1_fit, scaled_test_data)
test_data$Demand_pred_l2 <- predict(xgb_l2_fit, scaled_test_data)

cat("RMSE of L1 model is:",
     mean((test_data$Demand_pred_l1 - test_data$Demand)^2)^0.5,
    "\n")
cat("RMSE of L2 model is:",
     mean((test_data$Demand_pred_l2 - test_data$Demand)^2)^0.5,
    "\n")
```

This is quite a drop in performance. It's actually worse than what we see for some of the models in the previous doc, boosting-tests.Rmd.

When we plot the predictions we see there is quite a lot of variance.

```{r}
test_data %>%
  filter(Month == "Jun") %>% 
  plot_ly(x = ~ts) %>%
  add_trace(y = ~Demand, name = "Actual", type = "scatter", 
            mode = "lines") %>% 
  add_trace(y = ~Demand_pred_l1, name = "L1 regularization", type = "scatter", 
            mode = "lines") %>% 
  add_trace(y = ~Demand_pred_l2, name = "L2 regularization", type = "scatter", 
            mode = "lines") %>% 
  layout(title = "Actuals and model predictions")
```

