---
title: "Bootstrapping"
author: Cameron Roach
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
require(plotly)
require(DT)
require(caret)
require(myhelpr)
require(doMC)

registerDoMC(cores = 3)

source("../R/loadData.R")
source("../R/clean-smd-data.R")

load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)

smd <- load_smd_data(load_zones, root_dir = "./..")
smd <- clean_smd_data(smd, root_dir = "./..")

iZ <- "VT"

smd <- filter(smd, Zone == iZ)
```


# Bootstrapping methodology

Similar to the double seasonal block bootstrap by Hyndman and Shu (2010).

## Psuedo-code

The bootstrapping method will return date and time values so that any variable can be looked up later on. This will ensure any bootstrapped weather data will retain correlations between zones and with residuals.

The parameters `delta1` and `delta2` set the amount of variation in the date chosen and the length of the block sampled.

**NOTE: Be careful when sampling at end of year. Make sure we move from one year to the next and don't accidentally sample from beginning => discontinuity!**

```{}
GET days in season
SET delta1, delta2
SET number of simulated seasons

FOR number of simulated seasons
  WHILE simulated season not complete
    FOR current day of simulated season
      GET random historical year
      GET day near current day => current day + U(-delta1, delta1)
      GET the next 14 + U(-delta2, delta2) dates starting from this day
      APPEND these historical dates to current simulated season
      SET index to next day of simulated season without data.

GET weather data for dates and times in bootstrapped date-time data
GET residual data for dates and times in bootstrapped date-time data
```


## R function

```{r bootstrapping_function}
#' Double seasonal block bootstrap
#'
#' Generates bootstrapped dates using a double seasonal block bootstrap approach.
#'
#' @param date_series array of dates.
#' @param start_date start date for period we want bootstrap samples for.
#' @param end_date end date for period we want bootstrap samples for.
#' @param n_sim number of simulations. Defaults to 100 simulations.
#' @param avg_block_len average length of blocks. Defaults to 14 days.
#' @param delta_loc amount to randomise current location by when sampling blocks. See details.
#' @param delta_len amount to randomise block length by. See details.
#'
#' delta_loc shifts the current location in the simulated year by a random amount before sampling from a random historical year.
#'
#' @return A data frame with two columns. One for simulation number and another for bootstrapped dates.
#' @export
#'
#' @examples
#'
#' @author Cameron Roach
dbl_block_bs <- function(dates, start_date = today(), end_date = today() + 30, 
                         n_sim = 100, avg_block_len = 14, delta_loc = 0,
                         delta_len = 0) {
  
  dates <- sort(as.Date(unique(dates)))
  years <- unique(year(dates))
  end_date <- as.Date(end_date)
  start_date <- as.Date(start_date)
  bs_length <- end_date - start_date
  
  if (bs_length > 365) stop("Start and end dates should not be more than a year apart.")
  
  bs_sim <- NULL
  for (iSasdf in 1:n_sim) {
    # initialise
    bs_tmp <- NULL
    loc_date <- start_date
    while (loc_date < end_date) {
      year_sample <- sample(years, 1)
      block_loc <- loc_date + sample(-delta_loc:delta_loc, 1)
      block_length <- avg_block_len + sample(-delta_len:delta_len, 1)
      
      # update block location with sampled year
      year(block_loc) <- year_sample
      
      bs_dates <- block_loc + 0:(block_length - 1)
      
      # update
      bs_tmp <- bind_rows(bs_tmp,
                          data.frame(Simulation = iSasdf,
                                     Date = bs_dates))
      loc_date <- loc_date + block_length
    }
    
    bs_tmp <- bs_tmp[1:bs_length,]
    bs_sim <- bind_rows(bs_sim, bs_tmp)
  }
  
  return(bs_sim)
}
```




# Weather bootstrapping

Correlogram for dry bulb temperature and dew point temperature shown below. Serial correlations for 720 lags (30 days) have been calculated.

```{r weather_correlation}
with(smd, acf(DryBulb, lag.max = 30*24))

with(smd, acf(DewPnt, lag.max = 30*24))
```


```{r weather_bootstrapping}

```


# Residual bootstrapping
 
 
 
 ```{r residual_bootstrapping}

```
