---
title: "Bootstrapping"
author: "Cameron Roach"
output:
  pdf_document:
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
require(plotly)
require(DT)
require(caret)
require(myhelpr)
require(doMC)
require(gefcom2017)

registerDoMC(cores = 3)

load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)

smd <- load_smd_data(load_zones, root_dir = ".")
smd <- clean_smd_data(smd, root_dir = ".")

iZ <- "VT"

smd <- filter(smd, Zone == iZ)
```


# Bootstrapping methodology

Similar to the double seasonal block bootstrap by Hyndman and Shu (2010).

## Psuedo-code

The bootstrapping method will return date and time values so that any variable can be looked up later on. This will ensure any bootstrapped weather data will retain correlations between zones and with residuals.

The parameters `delta1` and `delta2` set the amount of variation in the date chosen and the length of the block sampled.

**NOTE: Be careful when sampling at end of year. Make sure we move from one year to the next and don't accidentally sample from beginning => discontinuity!**

```{}
GET days in season
SET delta1, delta2
SET number of simulated seasons

FOR number of simulated seasons
  WHILE simulated season not complete
    FOR current day of simulated season
      GET random historical year
      GET day near current day => current day + U(-delta1, delta1)
      GET the next 14 + U(-delta2, delta2) dates starting from this day
      APPEND these historical dates to current simulated season
      SET index to next day of simulated season without data.

GET weather data for dates and times in bootstrapped date-time data
GET residual data for dates and times in bootstrapped date-time data
```


## R function

```{r bootstrapping_function}
#' Double seasonal block bootstrap
#'
#' Generates bootstrapped dates using a double seasonal block bootstrap approach.
#'
#' @param date_series array of dates.
#' @param start_date start date for period we want bootstrap samples for.
#' @param end_date end date for period we want bootstrap samples for.
#' @param n_sims number of simulations. Defaults to 100 simulations.
#' @param avg_block_len average length of blocks. Defaults to 14 days.
#' @param delta_loc amount to randomise current location by when sampling blocks. See details.
#' @param delta_len amount to randomise block length by. See details.
#'
#' delta_loc shifts the current location in the simulated year by a random amount before sampling from a random historical year.
#'
#' @return A data frame with two columns. One for simulation number and another for bootstrapped dates.
#' @export
#'
#' @examples
#'
#' @author Cameron Roach
dbl_block_bs <- function(dates, start_date = today(), end_date = today() + 30, 
                         n_sims = 100, avg_block_len = 14, delta_loc = 3,
                         delta_len = 3) {
  
  dates <- sort(as.Date(unique(dates)))
  first_date <- min(dates)
  final_date <- max(dates)
  years <- unique(year(dates))
  end_date <- as.Date(end_date)
  start_date <- as.Date(start_date)
  bs_length <- end_date - start_date + 1
  date_seq <- seq(start_date, end_date, 1)
  if (start_date >= end_date) stop("Start date greater than or equal to end date.")
  if (bs_length > 365) stop("Start and end dates should not be more than a year apart.")
  
  bs_sim <- NULL
  for (iS in 1:n_sims) {
    # initialise
    bs_tmp <- NULL
    loc_date <- start_date
    while (loc_date <= end_date) {
      year_sample <- sample(years, 1)
      block_loc <- loc_date + sample(-delta_loc:delta_loc, 1)
      block_length <- avg_block_len + sample(-delta_len:delta_len, 1)
      
      # update block location with sampled year
      # resample if NA produced (leap years/start of dates cause issues)
      block_loc_bkp <- block_loc
      year(block_loc) <- year_sample
      if (is.na(block_loc) | block_loc < first_date) {
        block_loc <- block_loc_bkp
        next
      }
      
      bs_dates <- block_loc + 0:(block_length - 1)
      
      # resample if sampling dates after historical data ends
      if (tail(bs_dates, 1) > final_date) {
        block_loc <- block_loc_bkp
        next
      }
      # update
      bs_tmp <- bind_rows(bs_tmp,
                          data.frame(Simulation = iS,
                                     Date = bs_dates))
      loc_date <- loc_date + block_length
    }
    bs_tmp <- bs_tmp[1:bs_length,]
    bs_sim <- bind_rows(bs_sim, bs_tmp)
  }
  bs_sim$Date_seq <- rep(date_seq, n_sims)
  return(bs_sim)
}
```

Testing this function to make sure we have a uniform distribution of dates except near start and end date. There is a slight bump around 1st-4th March but this is likely due to resampling when we sample 29th February in a non-leap-year.

```{r test_bs_function}
#start_date = dmy("15/2/2016")
#end_date = dmy("18/3/2016")
start_date = dmy("1/10/2017")
end_date = dmy("1/4/2018")-1
# start_date = dmy("1/4/2017")
# end_date = dmy("1/10/2017")-1
n_sims = 1000
avg_block_len = 14
delta_loc = 3
delta_len = 3

bs_dates <- dbl_block_bs(smd$Date, start_date, end_date, n_sims, 
                         avg_block_len, delta_loc, delta_len)


table(ddmm = strftime(bs_dates$Date, "%m%d")) %>% 
  data.frame() %>% 
  mutate(ddmm = as.numeric(as.character(ddmm))) %>% 
  ggplot(aes(x=ddmm, y=Freq)) + 
  geom_col() + 
  ggtitle("Number of observations per ddmm",
          "ddmm used to see behaviour around 29/2")

table(Year = year(bs_dates$Date)) %>% 
  data.frame() %>% 
  ggplot(aes(x=Year, y=Freq)) +
  geom_col() +
  ggtitle("Number of observations per year.")
```



# Weather bootstrapping

Correlogram for dry bulb temperature and dew point temperature shown below. Serial correlations for 720 lags (30 days) have been calculated.

```{r weather_correlation}
with(smd, acf(DryBulb, lag.max = 30*24))

with(smd, acf(DewPnt, lag.max = 30*24))
```

I'm not seeing anything too meaningful there. We'll stick with 14 days for the average bootstrap length as per Hyndman & Shu (2010).

**TODO: Once we update the ts values, all the other calendar variables are messed up. Need to rerun each simulation through a create dummys function.**

```{r bootstrap_weather}
system.time({
  weather_bs <- smd %>% 
    mutate(Date = as.Date(Date)) %>% 
    right_join(bs_dates) %>% 
    mutate(ts = ymd_h(paste(Date_seq, Hour)))
})

# TODO: Now create dummy variables for each simulation.
```

Time-series plots of the dry bulb temperature are shown for the historical data and 16 of the bootstrapped simulations. Only February is plotted. The simulated values don't look unreasonable compared to actual historical values.

```{r}
plot_month = 2
weather_bs %>% 
  filter(month(ts) == plot_month,
         Simulation %in% sample(1:n_sims, 16)) %>% 
  ggplot(aes(x = ts, y = DryBulb)) + 
  geom_line(alpha = 0.3) +
  facet_wrap(~Simulation) +
  ggtitle("Simulated temperatures")

smd %>% 
  filter(month(ts) == plot_month) %>% 
  ggplot(aes(x = ts, y = DryBulb)) + 
  geom_line(alpha = 0.3) +
  facet_wrap(~Year, scales="free_x") +
  ggtitle("Historical temperatures")

```

Smoothing all of the historical years and the simulated years shows pretty similar paths. The historical year smoother has a lot more variance as there are only 13 years compared to 1000 for the simulated set.

```{r weather_bootstrapping}
ggplot() +
  geom_smooth(data = weather_bs %>% 
                mutate(DoY = yday(ts),
                       DoM = mday(ts),
                       Month = factor(month(ts, label=T)), ordered=F),
              aes(x = DoM, y = DryBulb, colour = "Simulation")) +
  geom_smooth(data = smd %>%
                filter(Month %in% unique(month(weather_bs$ts, label=T))) %>% 
                mutate(DoM = mday(ts)),
              aes(x = DoM, y = DryBulb, colour = "Historical")) +
  facet_wrap(~Month) +
  scale_colour_discrete(name = NULL) +
  ggtitle("Smoothed dry bulb temperature.")
```

Smoothed daily profiles are shown below for each month. Again, the historical and simulated values align quite well.

```{r}
ggplot() +
  geom_smooth(data = weather_bs %>% 
                mutate(Hour = hour(ts),
                       Month = factor(month(ts, label=T)), ordered=F),
              aes(x = Hour, y = DryBulb, colour = "Simulation")) +
  geom_smooth(data = smd %>% 
                filter(Month %in% unique(month(weather_bs$ts, label=T))),
              aes(x = Hour, y = DryBulb, colour = "Historical")) +
  facet_wrap(~Month) +
  scale_colour_discrete(name = NULL) +
  ggtitle("Smoothed daily dry bulb temperature")
```





# Residual bootstrapping

**IMPORTANT: Remember that residuals are only added to the peaks!!! Not the whole profile!!! Otherwise we would wind up with incredibly spiky profiles.**
 
## Fit model and calculate residuals

This needs to happen before the bootstrapping stage. Fit the best model and get residuals.

**TODO**

## Bootstrap residuals

Just use the same bootstrap dates as for the weather.

**TODO**
 
```{r residual_bootstrapping}

```

**TODO: Check how bootstrapped residuals look.**
