---
title: "Testing performance factors caret"
author: "Cameron Roach"
date: "30 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
require(plotly)
require(DT)
require(caret)
require(myhelpr)

source("../R/loadData.R")

load_zones_mass <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_mass)

smd <- load_smd_data(load_zones, root_dir = "./..")
```

Depending on the variable type caret can take a long time to fit xgboost. If a factor is ordered this severely impacts the run time. All other methods of structuring the training data are roughly the same speed with similar model performance.

It looks like the fit is slightly more accurate when creating my own dummy variables. Not sure why.

Seems to make sense to just keep everything in a data frame but make sure the factors aren't ordered.

```{r}
train_data <- filter(smd,
                     Year >= 2014,
                     Year < 2015,
                     Zone == "ME")
train_data_matrix <- data.matrix(train_data)

train_data$DoW_no <- factor(train_data$DoW, ordered = FALSE)
train_data$Month_no <- factor(train_data$Month, ordered = FALSE)

  
  
smd_fit <- list()

xgb_grid <- expand.grid(nrounds = 500,
                        eta = c(0.01,0.1),
                        max_depth = c(2,6,10),
                        gamma = 1,
                        colsample_bytree = 1,
                        min_child_weight = 1)


# user   system  elapsed 
# 1461.139   14.167 1503.170
system.time({
  smd_fit[[2]] <- train(Demand ~ DryBulb + DoW + Holiday_flag + Period + Month,
        data = train_data,
        method="xgbTree",
        metric="MAE",
        maximize = FALSE,
        tuneGrid = xgb_grid)
})

# user  system elapsed 
# 301.648   4.480 315.465 
system.time({
  smd_fit[[3]] <- train(Demand ~ DryBulb + DoW + Holiday_flag + Period + Month,
        data = train_data_matrix,
        method="xgbTree",
        metric="MAE",
        maximize = FALSE,
        tuneGrid = xgb_grid)
})


# Testing model 2 again to see if removing ordered factors improves performance 
# Looks like it is the ordering that messes things up. Caret is doing something
# extra under the hood.
# 
# user  system elapsed
# 323.105   5.181 342.080
system.time({
  smd_fit[[4]] <- train(Demand ~ DryBulb + DoW_no + Holiday_flag + Hour + Month_no,
                        data = train_data,
                        method="xgbTree",
                        metric="MAE",
                        maximize = FALSE,
                        tuneGrid = xgb_grid)
})


# user  system elapsed 
# 320.356   5.029 335.385
system.time({
  train_data_dummyVars <- dummyVars(~ Demand + DryBulb + DoW_no + Holiday_flag + Hour + Month_no,
                                    data = train_data,
                                    sep = "_",
                                    fullRank = TRUE)
  train_data_dummyVars <- predict(train_data_dummyVars, train_data)
  
  smd_fit[[5]] <- train(Demand ~ .,
                        data = train_data_dummyVars,
                        method="xgbTree",
                        metric="MAE",
                        maximize = FALSE,
                        tuneGrid = xgb_grid)
})

```

