---
title: "Monthly consumption forecasting."
author: "Cameron Roach"
output:
  pdf_document:
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = "png",
	dpi = 150
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
theme_set(theme_bw())
require(plotly)
require(caret)
require(myhelpr)
require(doMC)
require(gefcom2017)
require(WriteXLS)

registerDoMC(cores = 11)

# Inputs
load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)
agg_zones <- c("TOTAL", "MASS")
all_zones <- c("TOTAL", "ME", "NH", "VT", "CT", "RI", "MASS", load_zones_ma) # hierarchical order
hts_recon_flag <- TRUE
n_sims <- 100
fcst_start_date <- dmy("1/3/2015")
fcst_end_date <- dmy("1/3/2015") - 1

# Load data
smd <- load_smd_data(load_zones)
smd <- clean_smd_data(smd)

# separate data frames for aggregated zones because may change modelling, 
# e.g., remove average of variables and include all individual ones.
smd_mass <- smd %>% 
  filter(Zone %in% load_zones_ma) %>% 
  group_by(Date, Hour, Holiday, Holiday_flag, ts, Period, Year, Month, DoW,
           DoY, Weekend) %>% 
  summarise(Demand = sum(Demand),
            DryBulb = mean(DryBulb),
            DewPnt = mean(DewPnt),
            DryDewDiff = mean(DryDewDiff)) %>% 
  ungroup() %>% 
  mutate(Zone = "MASS")

smd_total <- smd %>% 
  group_by(Date, Hour, Holiday, Holiday_flag, ts, Period, Year, Month, DoW,
           DoY, Weekend) %>% 
  summarise(Demand = sum(Demand),
            DryBulb = mean(DryBulb),
            DewPnt = mean(DewPnt),
            DryDewDiff = mean(DryDewDiff)) %>% 
  ungroup() %>% 
  mutate(Zone = "TOTAL")

# create lagged predictors
smd <- smd %>%
  group_by(Zone) %>% 
  do(get_lagged_vars(., c("DryBulb", "DewPnt"), lags = 1:72)) %>% 
  ungroup() %>% 
  filter(Year >= 2009)

smd_mass <- smd_mass %>% 
  do(get_lagged_vars(., c("DryBulb", "DewPnt"), lags = 1:72)) %>% 
  filter(Year >= 2009)

smd_total <- smd_total %>% 
  do(get_lagged_vars(., c("DryBulb", "DewPnt"), lags = 1:72)) %>% 
  filter(Year >= 2009)
```


# Question

Can we predict monthly demand using univariate time-series techniques? If we can we may be able to improve the model performance.

# Data

```{r}
smd %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(Zone, MMYY) %>% 
  summarise(Average_demand = mean(Demand)) %>% 
  ggplot(aes(x = MMYY, y = Average_demand)) + 
  geom_line() + 
  facet_wrap(~Zone, scales = "free_y") +
  ggtitle("Average demand by load zone")

smd_mass %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(MMYY) %>% 
  summarise(Average_demand = mean(Demand)) %>% 
  ggplot(aes(x = MMYY, y = Average_demand)) + 
  geom_line() + 
  ggtitle("Average demand in Massachusetts")

smd_total %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(MMYY) %>% 
  summarise(Average_demand = mean(Demand)) %>% 
  ggplot(aes(x = MMYY, y = Average_demand)) + 
  geom_line() + 
  ggtitle("Average demand for total of all zones")
```

```{r}
smd %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(Month, Zone, MMYY) %>% 
  summarise(Average_demand = mean(Demand)) %>% 
  ggplot(aes(x = MMYY, y = Average_demand, colour = Zone)) + 
  geom_line() + 
  geom_smooth(method = "lm") +
  facet_wrap(~Month) +
  ggtitle("Average demand by load zone")

smd_mass %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(Month, MMYY) %>% 
  summarise(Average_demand = mean(Demand)) %>% 
  ggplot(aes(x = MMYY, y = Average_demand)) + 
  geom_line() + 
  geom_smooth(method = "lm") +
  facet_wrap(~Month) + 
  ggtitle("Average demand in Massachusetts")

smd_total %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(Month, MMYY) %>% 
  summarise(Average_demand = mean(Demand)) %>% 
  ggplot(aes(x = MMYY, y = Average_demand)) + 
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~Month) + 
  ggtitle("Average demand for total of all zones")
```


# Two-model approach

1. Instead of forecasting quantiles of demand, forecast quantiles of normalised demand.
2. Using univariate time-series technique, forecast consumption for coming month.
3. Scale quantiles by the forecast monthly consumption.

```{r arima_forecast}
require(forecast)
require(ggplot2)
require(scales)

smd_total_avg <- smd_total %>%
  mutate(MMYY = floor_date(ts, "month")) %>% 
  group_by(MMYY) %>% 
  summarise(Average_demand = mean(Demand))

y <- ts(smd_total_avg$Average_demand, 
        freq=12,
        start=c(2003, 3))
yt <- window(y,end=2015)
#auto.arima(yt)
yfit <- Arima(yt, order=c(1,0,0), seasonal=c(2,1,1), lambda=0)
yfor <- forecast(yfit)
#pd <- funggcast(y,yfor)

autoplot(yfor)+ 
  geom_line(data = data.frame(Date = time(y),
                              Demand = as.numeric(y)),
            aes(x=Date, y=Demand),
            colour = "black")
```



# Background

Have noticed (only ad-hoc tests) that in a year of high consumption all the quantile forecasts seem to be negatively biased (actuals are always closer to the 90th quantile) and vice versa for low consumption years (actuals usually near 10th quantile).

Need to investigate how we can go about forecasting monthly demand and how confident we can be.

## Added uncertainty

We need to take the uncertainty in this monthly forecast into consideration. Can use prediction intervals.

**TODO: Need to figure out how to incorporate.**

# Single model approach

Instead of using two models, it might be possible to just use average monthly demand in last two-three months as input variables. These values won't be bootstrapped as we should have actual values available.
