---
title: "Variable selection in aggregated zones"
author: "Cameron Roach"
date: "29 January 2017"
output:
  pdf_document:
    includes:
      in_header: mystyles.sty
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = "png",
	dpi = 150
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
theme_set(theme_bw())
require(plotly)
require(caret)
require(myhelpr)
require(doMC)
require(gefcom2017)

registerDoMC(cores = 11)

# Inputs
load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)
agg_zones <- c("MASS", "TOTAL")
all_zones <- c(load_zones, agg_zones)
n_sims <- 1000
fcst_start_date <- dmy("1/2/2017")
fcst_end_date <- dmy("1/3/2017") - 1

# Load data
smd <- load_smd_data(load_zones, root_dir = ".")
smd <- clean_smd_data(smd, root_dir = ".")

# Sort out MASS and TOTAL data frames. No lags.
smd_mass_avg <- smd %>% 
  filter(Zone %in% load_zones_ma) %>% 
  group_by(Date, Hour, Holiday, Holiday_flag, ts, Period, Year, Month, DoW,
           DoY, Weekend) %>% 
  summarise(Demand = sum(Demand),
            DryBulb = mean(DryBulb),
            DewPnt = mean(DewPnt),
            DryDewDiff = mean(DryDewDiff)) %>% 
  ungroup() %>% 
  mutate(Zone = "MASS")

smd_mass_all <- smd %>% 
    filter(Zone %in% load_zones_ma) %>% 
    gather(Var, Val, c(DryBulb, DewPnt, DryDewDiff, Demand)) %>% 
    unite(Zone_Var, Zone, Var) %>% 
    spread(Zone_Var, Val)
smd_mass_all$Demand <- smd_mass_all %>% 
  select(contains("_Demand")) %>% 
  rowSums()
smd_mass_all <- smd_mass_all %>% 
  select(-contains("_Demand")) %>% 
  mutate(Zone = "MASS")

smd_total_avg <- smd %>% 
  group_by(Date, Hour, Holiday, Holiday_flag, ts, Period, Year, Month, DoW,
           DoY, Weekend) %>% 
  summarise(Demand = sum(Demand),
            DryBulb = mean(DryBulb),
            DewPnt = mean(DewPnt),
            DryDewDiff = mean(DryDewDiff)) %>% 
  ungroup() %>% 
  mutate(Zone = "TOTAL")

smd_total_all <- smd %>% 
    gather(Var, Val, c(DryBulb, DewPnt, DryDewDiff, Demand)) %>% 
    unite(Zone_Var, Zone, Var) %>% 
    spread(Zone_Var, Val)
smd_total_all$Demand <- smd_total_all %>% 
  select(contains("_Demand")) %>% 
  rowSums()
smd_total_all <- smd_total_all %>% 
  select(-contains("_Demand")) %>% 
  mutate(Zone = "MASS")
```


# Introduction

Here we test variable selection methods for aggregated zones TOTAL and MASS against a baseline model. The variable selection methods include L1 and L2 penalisation within a linear booster. The baseline model will simply be the average of all weather stations that fall within an aggregated zone. The baseline model will also be tested with L1 and L2 regularization to ensure consistency.

First we will test an underspecified model that does not include lags. We will conclude by testing a full specified model. The baseline model will be penalised in this case to ensure that only relevant lags are being included in the final model.


# Model fitting - no lags

Using the years 2009 to 2015 as the training set. 2016 will be omitted and used later as the test set.

```{r fit_model_no_lags}
xgb_ctrl <- trainControl(method = "repeatedcv",
                         number = 5,
                         allowParallel = TRUE)


if (file.exists("./cache/variable-selection-in-aggregated-zones-no-lags.RData")) {
  load("./cache/variable-selection-in-aggregated-zones-no-lags.RData")
} else {
  system.time({
    xgb_fit <- NULL
    
    cat("Fitting TOTAL with averaged variables. L1 penalty...\n")
    xgb_fit[["TOTAL"]][["Averaged-variables-L1-regularization"]] <- smd_total_avg %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with averaged variables. L2 penalty...\n")
    xgb_fit[["TOTAL"]][["Averaged-variables-L2-regularization"]] <- smd_total_avg %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with all variables. L1 penalty...\n")
    xgb_fit[["TOTAL"]][["All-variables-L1-regularization"]] <- smd_total_all %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with all variables. L2 penalty...\n")
    xgb_fit[["TOTAL"]][["All-variables-L2-regularization"]] <- smd_total_all %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
  })
  
  save(xgb_fit, file = "./cache/variable-selection-in-aggregated-zones-no-lags.RData")
}
```

## CV performance

```{r cv_performance}
results_df <- NULL
for (iZ in names(xgb_fit)) {
  for (iM in names(xgb_fit[[iZ]])) {
    results_df <- data.frame(Zone = iZ,
                             Model = iM,
                             xgb_fit[[iZ]][[iM]]$results) %>% 
      bind_rows(results_df, .)
  }
}

ggplot() +
  geom_line(data = filter(results_df, Model == "Averaged-variables-L1-regularization"), 
            aes(x = alpha, y = RMSE, colour = "L1-regularization", linetype = "Averaged-variables")) +
  geom_line(data = filter(results_df, Model == "Averaged-variables-L2-regularization"), 
            aes(x = lambda, y = RMSE, colour = "L2-regularization", linetype = "Averaged-variables")) +
  geom_line(data = filter(results_df, Model == "All-variables-L1-regularization"), 
            aes(x = alpha, y = RMSE, colour = "L1-regularization", linetype = "All-variables")) +
  geom_line(data = filter(results_df, Model == "All-variables-L2-regularization"), 
            aes(x = lambda, y = RMSE, colour = "L2-regularization", linetype = "All-variables")) +
  ggtitle("5-fold CV RMSE by zone",
          "No lagged variables included") +
  xlab("Penalty") +
  scale_x_continuous(trans = "log",
                     labels = function(x) round(x, 0)) +
  labs(linetype = "Variables",
       colour = "Regularization") +
  facet_wrap(~Zone, scales = "free_y", nrow = 2)
```
