---
title: "Variable selection in aggregated zones"
author: "Cameron Roach"
date: "29 January 2017"
output:
  pdf_document:
    includes:
      in_header: mystyles.sty
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = "png",
	dpi = 150
)

rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
theme_set(theme_bw())
require(plotly)
require(caret)
require(myhelpr)
require(doMC)
require(gefcom2017)

registerDoMC(cores = 11)

# Inputs
load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)
agg_zones <- c("MASS", "TOTAL")
all_zones <- c(load_zones, agg_zones)
n_sims <- 1000
fcst_start_date <- dmy("1/2/2017")
fcst_end_date <- dmy("1/3/2017") - 1

# Load data
smd <- load_smd_data(load_zones, root_dir = ".")
smd <- clean_smd_data(smd, root_dir = ".")

# Sort out MASS and TOTAL data frames. No lags.
smd_mass_avg <- smd %>% 
  filter(Zone %in% load_zones_ma) %>% 
  group_by(Date, Hour, Holiday, Holiday_flag, ts, Period, Year, Month, DoW,
           DoY, Weekend) %>% 
  summarise(Demand = sum(Demand),
            DryBulb = mean(DryBulb),
            DewPnt = mean(DewPnt),
            DryDewDiff = mean(DryDewDiff)) %>% 
  ungroup() %>% 
  mutate(Zone = "MASS")

smd_mass_all <- smd %>% 
    filter(Zone %in% load_zones_ma) %>% 
    gather(Var, Val, c(DryBulb, DewPnt, DryDewDiff, Demand)) %>% 
    unite(Zone_Var, Zone, Var) %>% 
    spread(Zone_Var, Val)
smd_mass_all$Demand <- smd_mass_all %>% 
  select(contains("_Demand")) %>% 
  rowSums()
smd_mass_all <- smd_mass_all %>% 
  select(-contains("_Demand")) %>% 
  mutate(Zone = "MASS")

smd_total_avg <- smd %>% 
  group_by(Date, Hour, Holiday, Holiday_flag, ts, Period, Year, Month, DoW,
           DoY, Weekend) %>% 
  summarise(Demand = sum(Demand),
            DryBulb = mean(DryBulb),
            DewPnt = mean(DewPnt),
            DryDewDiff = mean(DryDewDiff)) %>% 
  ungroup() %>% 
  mutate(Zone = "TOTAL")

smd_total_all <- smd %>% 
    gather(Var, Val, c(DryBulb, DewPnt, DryDewDiff, Demand)) %>% 
    unite(Zone_Var, Zone, Var) %>% 
    spread(Zone_Var, Val)
smd_total_all$Demand <- smd_total_all %>% 
  select(contains("_Demand")) %>% 
  rowSums()
smd_total_all <- smd_total_all %>% 
  select(-contains("_Demand")) %>% 
  mutate(Zone = "MASS")
```


# Introduction

Here we test variable selection methods for aggregated zones TOTAL and MASS against a baseline model. The variable selection methods include L1 and L2 penalisation within a linear booster. The baseline model will simply be the average of all weather stations that fall within an aggregated zone. The baseline model will also be tested with L1 and L2 regularization to ensure consistency.

First we will test an underspecified model that does not include lags. We will conclude by testing a full specified model. The baseline model will be penalised in this case to ensure that only relevant lags are being included in the final model.


# Model evaluation - no lags

## Data and model fitting

Data in the years 2009 to 2015 will be used as the training set. 2016 will be omitted and used later as the test set.

```{r fit_model_no_lags}
xgb_ctrl <- trainControl(method = "repeatedcv",
                         number = 5,
                         allowParallel = TRUE)


if (file.exists("./cache/variable-selection-in-aggregated-zones-no-lags.RData")) {
  load("./cache/variable-selection-in-aggregated-zones-no-lags.RData")
} else {
  system.time({
    xgb_fit <- NULL
    
    cat("Fitting TOTAL with averaged variables. L1 penalty...\n")
    xgb_fit[["TOTAL"]][["Averaged-variables-L1-regularization"]] <- smd_total_avg %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with averaged variables. L2 penalty...\n")
    xgb_fit[["TOTAL"]][["Averaged-variables-L2-regularization"]] <- smd_total_avg %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with all variables. L1 penalty...\n")
    xgb_fit[["TOTAL"]][["All-variables-L1-regularization"]] <- smd_total_all %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with all variables. L2 penalty...\n")
    xgb_fit[["TOTAL"]][["All-variables-L2-regularization"]] <- smd_total_all %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with averaged variables. L1 penalty...\n")
    xgb_fit[["MASS"]][["Averaged-variables-L1-regularization"]] <- smd_mass_avg %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with averaged variables. L2 penalty...\n")
    xgb_fit[["MASS"]][["Averaged-variables-L2-regularization"]] <- smd_mass_avg %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with all variables. L1 penalty...\n")
    xgb_fit[["MASS"]][["All-variables-L1-regularization"]] <- smd_mass_all %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with all variables. L2 penalty...\n")
    xgb_fit[["MASS"]][["All-variables-L2-regularization"]] <- smd_mass_all %>% 
      filter(Year %in% 2009:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
  })
  
  save(xgb_fit, file = "./cache/variable-selection-in-aggregated-zones-no-lags.RData")
}
```

## CV performance

```{r cv_performance_no_lags}
results_df <- NULL
for (iZ in names(xgb_fit)) {
  for (iM in names(xgb_fit[[iZ]])) {
    results_df <- data.frame(Zone = iZ,
                             Model = iM,
                             xgb_fit[[iZ]][[iM]]$results) %>% 
      bind_rows(results_df, .)
  }
}

ggplot() +
  geom_line(data = filter(results_df, Model == "Averaged-variables-L1-regularization"), 
            aes(x = alpha, y = RMSE, colour = "L1-regularization", linetype = "Averaged-variables")) +
  geom_line(data = filter(results_df, Model == "Averaged-variables-L2-regularization"), 
            aes(x = lambda, y = RMSE, colour = "L2-regularization", linetype = "Averaged-variables")) +
  geom_line(data = filter(results_df, Model == "All-variables-L1-regularization"), 
            aes(x = alpha, y = RMSE, colour = "L1-regularization", linetype = "All-variables")) +
  geom_line(data = filter(results_df, Model == "All-variables-L2-regularization"), 
            aes(x = lambda, y = RMSE, colour = "L2-regularization", linetype = "All-variables")) +
  ggtitle("5-fold CV RMSE by zone",
          "No lagged variables included") +
  xlab("Penalty") +
  scale_x_continuous(trans = "log",
                     labels = function(x) round(x, 0)) +
  labs(linetype = "Variables",
       colour = "Regularization") +
  facet_wrap(~Zone, scales = "free_y")
```



## Test data results

```{r test_performance_no_lags}
df_lkp <- data_frame(
  Zone = rep(c("TOTAL", "MASS"), each = 4),
  Model = rep(c("Averaged-variables-L1-regularization",
                "Averaged-variables-L2-regularization",
                "All-variables-L1-regularization",
                "All-variables-L2-regularization"), 2),
  DataFrameName = rep(c("smd_total_avg", "smd_total_all", "smd_mass_avg", 
                         "smd_mass_all"), each = 2)
)

test_results_df <- NULL
for (iZ in names(xgb_fit)) {
  for (iM in names(xgb_fit[[iZ]])) {
    df_name <- df_lkp %>% 
      filter(Zone == iZ,
             Model == iM) %>% 
      .$DataFrameName
    
    test_data <- get(df_name) %>% 
      filter(Year == 2016)
    
    test_results_df <- data.frame(
      Zone = iZ,
      Model = iM,
      Demand = test_data$Demand,
      Demand_pred = predict(xgb_fit[[iZ]][[iM]], test_data)) %>% 
      bind_rows(test_results_df, .)
  }
}

test_rmse <- test_results_df %>% 
  group_by(Zone, Model) %>% 
  summarise(rmse = mean((Demand_pred - Demand)^2)^0.5) %>% 
  ungroup()

test_rmse %>% 
  ggplot(aes(x = Model, y = rmse)) +
  geom_point(size = 2) +
  ggtitle("RMSE for models on test data set") +
  coord_flip() +
  facet_wrap(~Zone, scales = "free_x", ncol = 1)
```


# Model evaluation - with lags

We will now repeat this simple analysis for a fully specified model containing all lags for the temperature variables.

## Data and model fitting

Lagged dummy variables are created for the dry bulb and dew point temperature variables.

```{r create_lags}
smd_mass_avg <- smd_mass_avg %>%
  do(get_lagged_vars(., 
                     names(select(., contains("DryBulb"), contains("DewPnt"))),
                     lags = 1:72))

smd_mass_all <- smd_mass_all %>%
  do(get_lagged_vars(., 
                     names(select(., contains("DryBulb"), contains("DewPnt"))),
                     lags = 1:72))

smd_total_avg <- smd_total_avg %>%
  do(get_lagged_vars(., 
                     names(select(., contains("DryBulb"), contains("DewPnt"))),
                     lags = 1:72))

smd_total_all <- smd_total_all %>%
  do(get_lagged_vars(., 
                     names(select(., contains("DryBulb"), contains("DewPnt"))),
                     lags = 1:72))
```

Once again, data in the years 2009 to 2015 will be used as the training set. 2016 will be omitted and used later as the test set.

```{r fit_model_lags}
# TODO: This hangs sometimes (11 threads but 0% CPU). Runs if setup chunk and create_lags chunk are the only chunks run prior. Try and include more chunks and see which one causes this to hang. Work out bug.
# Note: I'm only fitting 2012 to 2015 data here - had to cut down training data size to get it to work :(

xgb_ctrl <- trainControl(method = "repeatedcv",
                         number = 5,
                         allowParallel = TRUE)


if (file.exists("./cache/variable-selection-in-aggregated-zones-lags.RData")) {
  load("./cache/variable-selection-in-aggregated-zones-lags.RData")
} else {
  system.time({
    xgb_fit_lags <- NULL
    
    cat("Fitting TOTAL with averaged variables. L1 penalty...\n")
    xgb_fit_lags[["TOTAL"]][["Averaged-variables-L1-regularization"]] <- smd_total_avg %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with averaged variables. L2 penalty...\n")
    xgb_fit_lags[["TOTAL"]][["Averaged-variables-L2-regularization"]] <- smd_total_avg %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with all variables. L1 penalty...\n")
    xgb_fit_lags[["TOTAL"]][["All-variables-L1-regularization"]] <- smd_total_all %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting TOTAL with all variables. L2 penalty...\n")
    xgb_fit_lags[["TOTAL"]][["All-variables-L2-regularization"]] <- smd_total_all %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with averaged variables. L1 penalty...\n")
    xgb_fit_lags[["MASS"]][["Averaged-variables-L1-regularization"]] <- smd_mass_avg %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with averaged variables. L2 penalty...\n")
    xgb_fit_lags[["MASS"]][["Averaged-variables-L2-regularization"]] <- smd_mass_avg %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with all variables. L1 penalty...\n")
    xgb_fit_lags[["MASS"]][["All-variables-L1-regularization"]] <- smd_mass_all %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = c(0, exp(0:8)),
                                   lambda = 0,
                                   eta = 0.1),
            nthread = 1)
    
    cat("Fitting MASS with all variables. L2 penalty...\n")
    xgb_fit_lags[["MASS"]][["All-variables-L2-regularization"]] <- smd_mass_all %>% 
      #filter(Year %in% 2009:2015) %>% 
      filter(Year %in% 2012:2015) %>% 
      select(Demand, Hour, DoY, DoW, Holiday_flag, contains("DryBulb"), 
             contains("DewPnt")) %>%
      train(Demand ~ . ,
            data = .,
            method="xgbLinear",
            trControl = xgb_ctrl,
            tuneGrid = expand.grid(nrounds = 300,
                                   alpha = 0,
                                   lambda = c(0, exp(0:8)),
                                   eta = 0.1),
            nthread = 1)
    
  })
  
  save(xgb_fit_lags, file = "./cache/variable-selection-in-aggregated-zones-lags.RData")
}
```


```{r cv_performance_lags}
results_df_lags <- NULL
for (iZ in names(xgb_fit_lags)) {
  for (iM in names(xgb_fit_lags[[iZ]])) {
    results_df_lags <- data.frame(Zone = iZ,
                             Model = iM,
                             xgb_fit_lags[[iZ]][[iM]]$results) %>% 
      bind_rows(results_df_lags, .)
  }
}

ggplot() +
  geom_line(data = filter(results_df_lags, Model == "Averaged-variables-L1-regularization"), 
            aes(x = alpha, y = RMSE, colour = "L1-regularization", linetype = "Averaged-variables")) +
  geom_line(data = filter(results_df_lags, Model == "Averaged-variables-L2-regularization"), 
            aes(x = lambda, y = RMSE, colour = "L2-regularization", linetype = "Averaged-variables")) +
  geom_line(data = filter(results_df_lags, Model == "All-variables-L1-regularization"), 
            aes(x = alpha, y = RMSE, colour = "L1-regularization", linetype = "All-variables")) +
  geom_line(data = filter(results_df_lags, Model == "All-variables-L2-regularization"), 
            aes(x = lambda, y = RMSE, colour = "L2-regularization", linetype = "All-variables")) +
  ggtitle("5-fold CV RMSE by zone",
          "Lagged variables included") +
  xlab("Penalty") +
  scale_x_continuous(trans = "log",
                     labels = function(x) round(x, 0)) +
  labs(linetype = "Variables",
       colour = "Regularization") +
  facet_wrap(~Zone, scales = "free_y")
```


## Test data results

```{r test_performance_lags}
df_lkp <- data_frame(
  Zone = rep(c("TOTAL", "MASS"), each = 4),
  Model = rep(c("Averaged-variables-L1-regularization",
                "Averaged-variables-L2-regularization",
                "All-variables-L1-regularization",
                "All-variables-L2-regularization"), 2),
  DataFrameName = rep(c("smd_total_avg", "smd_total_all", "smd_mass_avg", 
                         "smd_mass_all"), each = 2)
)

test_results_df_lags <- NULL
for (iZ in names(xgb_fit_lags)) {
  for (iM in names(xgb_fit_lags[[iZ]])) {
    df_name <- df_lkp %>% 
      filter(Zone == iZ,
             Model == iM) %>% 
      .$DataFrameName
    
    test_data <- get(df_name) %>% 
      filter(Year == 2016)
    
    test_results_df_lags <- data.frame(
      Zone = iZ,
      Model = iM,
      Demand = test_data$Demand,
      Demand_pred = predict(xgb_fit_lags[[iZ]][[iM]], test_data)) %>% 
      bind_rows(test_results_df_lags, .)
  }
}

test_rmse_lags <- test_results_df_lags %>% 
  group_by(Zone, Model) %>% 
  summarise(rmse = mean((Demand_pred - Demand)^2)^0.5) %>% 
  ungroup()

test_rmse_lags %>% 
  ggplot(aes(x = Model, y = rmse)) +
  geom_point(size = 2) +
  ggtitle("RMSE for models on test data set") +
  coord_flip() +
  facet_wrap(~Zone, scales = "free_x", ncol = 1)
```
