---
title: "SMD exploration 1"
author: Cameron Roach
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
rm(list=ls())

require(dplyr)
require(tidyr)
require(readxl)
require(lubridate)
require(ggplot2)
require(plotly)
require(caret)
require(myhelpr)

source("../R/loadData.R")

load_zones_ma <- c("SEMASS", "WCMASS", "NEMASSBOST")
load_zones <- c("ME", "NH", "VT", "CT", "RI", load_zones_ma)

smd <- load_smd_data(root_dir = "./..", load_zones)
```

# Introduction

This document shows some interesting plots of the data. Also some dumb plots.


# Weather variables

## Temperature and demand

```{r}
smd %>% 
  filter(Year == 2015) %>% 
  ggplot(aes(x = DryBulb, y = Demand, colour = Weekend)) +
  geom_point(shape = 21, alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~Zone) +
  ggtitle("Demand and dry bulb temperature")

smd %>% 
  filter(Year == 2015) %>% 
  ggplot(aes(x = DewPnt, y = Demand, colour = Weekend)) +
  geom_point(shape = 21, alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~Zone) +
  ggtitle("Demand and dew point temperature")

smd %>% 
  filter(Year == 2015) %>% 
  ggplot(aes(x = DewPnt, y = DryBulb, colour = Weekend)) +
  geom_point(shape = 21, alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~Zone) +
  ggtitle("Dry bulb and dew point temperature")
```


## Differencing between dry bulb and dew point

```{r}
smd %>% 
  filter(Year == 2015) %>% 
  select(Zone, DewPnt, DryBulb, DryDewDiff) %>% 
  gather(TempType, Temp, -c(DryDewDiff, Zone)) %>% 
  ggplot(aes(x = DryDewDiff, y = Temp, colour=TempType)) +
  geom_point(shape = 21, alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~Zone) +
  ggtitle("Dry bulb and dew point difference correlations")

smd %>% 
  filter(Year == 2015) %>% 
  ggplot(aes(x = DryDewDiff, y = Demand, colour = Weekend)) +
  geom_point(shape = 21, alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~Zone, scale = "free_y") +
  ggtitle("Demand and difference betweeen dry bulb and dew point temperature")
```

## Day of week, temperature and demand

I'm pretty sure this plot is dumb.

```{r}
smd %>% 
  filter(Year == 2015) %>% 
  ggplot(aes(x = DryBulb, y = Demand, colour = DoW)) +
  #geom_point(shape = 21, alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_wrap(~Zone, scales = "free_y") +
  ggtitle("Demand and dry bulb temperature")
```


# Demand

## Demand outlier checks

We can see in these plots that there are some clear outliers in the latest data file for 2016 (surprise!).

```{r}
smd %>% 
  group_by(Zone, Date) %>% 
  summarise(Demand = sum(Demand)) %>% 
  ungroup() %>% 
  ggplot(aes(x=Date, y=Demand)) +
  geom_line() +
  facet_wrap(~Zone) + 
  ggtitle("Daily demand in each zone")

smd %>% 
  ggplot(aes(x=Month, y=Demand, colour=Month)) + 
  geom_boxplot() +
  facet_wrap(~Zone, scales = "free_y") +
  ggtitle("Boxplots of demand in each zone by month")
```


## Averaged profiles

First plot shows average across the whole week. Second plot shows average for each day.

```{r}
smd %>% 
  group_by(Zone, Hour) %>% 
  summarise(Demand = mean(Demand)) %>% 
  ggplot(aes(x=Hour, y=Demand)) +
  geom_line() +
  facet_wrap(~Zone) +
  ggtitle("Average daily demand profiles")

smd %>% 
  group_by(Zone, Hour, DoW) %>% 
  summarise(Demand = mean(Demand)) %>% 
  ggplot(aes(x=Hour, y=Demand, colour=DoW)) +
  geom_line() +
  facet_wrap(~Zone, scales="free_y") +
  ggtitle("Average daily demand profiles")
```

## Daylight savings impact

Is there any evidence of daylight savings changing demand behaviour from month to month? Want to check if ramp up in morning and decrease at night are still happening in the same place. Can't compare peak as there is clear evidence of solar (which pushes the peak  back).

```{r}
for (iZ in load_zones) {
  cc <- scales::seq_gradient_pal("blue", "red", "Lab")(seq(0,1,length.out=12))

  p <- smd %>% 
    filter(Zone == iZ) %>% 
    group_by(Month, Hour) %>% 
    summarise(Demand = mean(Demand)) %>% 
    ggplot(aes(x=Hour, y=Demand, colour=Month)) +
    geom_line() +
    scale_colour_manual(values=cc) +
    ggtitle(paste("Average daily demand profiles by month for", iZ))
  print(p)
}
```


## Demand comparison between months

```{r}
for (iZ in load_zones) {
  cc <- scales::seq_gradient_pal("blue", "red", "Lab")(seq(0,1,length.out=12))

  p <- smd %>% 
    filter(Zone == iZ) %>% 
    group_by(Month, Hour) %>% 
    summarise(Demand = mean(Demand)) %>% 
    ggplot(aes(x=Hour, y=Demand, colour=Month)) +
    geom_line() +
    scale_colour_manual(values=cc) +
    facet_wrap(~Month) +
    ggtitle(paste("Average daily demand profiles by month for", iZ))
  print(p)
}
```

## Empirical percentiles of demand profiles

```{r}
for (iZ in load_zones) {
  quantiles <- 1:9/10
  p <- smd %>% 
    filter(Zone == iZ) %>% 
    group_by(Hour, Month) %>% 
    na.omit() %>% 
    do(data_frame(
      Demand = quantile(na.omit(.$Demand), quantiles),
      Quantile = quantiles)) %>% 
    ggplot(aes(x=Hour, y=Demand, colour=Quantile, group=Quantile)) +
    geom_line() +
    facet_wrap(~Month) +
    ggtitle(paste("Average daily demand profiles by month for", iZ))
  print(p)
}
```


# Public holidays

It appears as though some public holidays have a bigger impact on demand than others (Christmas Day, Independence day). Need to investigate if this is because of other factors such as falling after a weekend or interaction with holiday periods.

Individual boxplots were also produced for each load zone but gave very similar results.

** TODO: How do public holidays compare to weekend days? Does this mean that some public holidays can be modelled as weekends (reduces dimensionality)? **


```{r}
smd %>% 
  group_by(Month, Date, Holiday, Holiday_flag) %>% 
  summarise(Demand = sum(Demand)) %>% 
  group_by(Month) %>% 
  mutate(Demand = Demand/mean(Demand)) %>% 
  filter(Holiday_flag == TRUE) %>% 
  ggplot(aes(x=Holiday, y=Demand)) +
  geom_boxplot() +
  ggtitle(paste("Normalised demand on public holidays"))
```

